DBES Grand Challenge : High throughput, scalable solution for processing taxi trip data

### 2015-03-24 paper/experiments.tex
fixing typos

\section{Experiments}
We conduced several experiments to measure the soundness of our solution. Since the parallel message processing gives the better performance, we present details of those experiments with distributed setup as well a singel JVM setup. All these experiments were performed on a LAN with a network bandwidth of 1Gbps and all machines were Intel(R) Xeon(R) 2.4GHz, 4 core duo machines with 16 GB of memory.

@@

We deployed this process graph to a distributed stream processing system we have developed, and  measured the execution time and the average message latency varying the number of route processors with the data set given (173 million records). In local set up, we use a new thread to execute a new instance and in distributed setup we add a new node to execute the new instance.

@@

Figure \ref{latency_route} shows the latencies. For single JVM setup latency get reduced at 2 and slightly get increase due to context switching as in the execution times. For distributed set up initial latency is high due to communication and message serialization parsing overheads. With higher number of processors latency get reduced.

@@

Figure \ref{profictgraph} shows the process graph for profitable cells query. \textit{ProfitEventEmitter} reads the data file and calculates the pickup cell and dropoff cell details using longitude and latitude coordinates. Then it creates two events one for dropoff part and other to pickup part, with different sequence numbers and distributes the events among \textit{ProfictCalculators} using the hash value of the cell. In this way cells are partitioned into different \textit{ProfitCalculator} instances and each instance gets all the events relevant to its' cell set. We use a set of threads (equals to the number of Profict Calculators) to emit data as in the previous query. Finally \textit{TopProfictProcessor} aggregates events and writes to a file. 
Similar to earlier experiment,  we deployed this process graph to our system and measured the execution time and the average message latency with the data set given for both single JVM setup and distributed setup.

>>>>

\section{Experiments}
We conduced several experiments to measure the soundness of our solution. Since the parallel message processing gives the better performance, we present details of those experiments with distributed setup as well a single JVM setup. All these experiments were performed on a LAN with a network bandwidth of 1Gbps and all machines were Intel(R) Xeon(R) 2.4GHz, 4 core duo machines with 16 GB of memory.

@@

We deployed this process graph to a distributed stream processing system we have developed, and  measured the execution time and the average message latency varying the number of route processors with the data set given (173 million records). In the local set up, we use a new thread to execute a new instance and in the distributed setup we add a new node to execute the new instance.

@@

Figure \ref{latency_route} shows the latencies. For single JVM setup latency get reduced at 2 and slightly increases due to context switching as in the execution times. For distributed set up initial latency is high due to communication and message serialization parsing overheads. With higher number of processors latency gets reduced.

@@

Figure \ref{profictgraph} shows the process graph for profitable cells query. \textit{ProfitEventEmitter} reads the data file and calculates the pickup cell and dropoff cell details using longitude and latitude coordinates. Then it creates two events one for dropoff part and other to pickup part, with different sequence numbers and distributes the events among \textit{ProfitCalculators} using the hash value of the cell. In this way cells are partitioned into different \textit{ProfitCalculator} instances and each instance gets all the events relevant to its' cell set. We use a set of threads (equals to the number of Profict Calculators) to emit data as in the previous query. Finally \textit{TopProfitProcessor} aggregates events and writes to a file. 
Similar to earlier experiment,  we deployed this process graph to our system and measured the execution time and the average message latency with the data set given for both single JVM setup and distributed setup.

### 2015-03-24 paper/paper.tex
updated the paper
fixing typos

\maketitle
\begin{abstract}
Processing complex queries on unbounded event streams in real time, is a challenge for many organizational data processing systems. These systems are expected to process data with minimal latency to generate real time events, and high through put to minimize the required hardware. In this regard DEBS Grand Challenge 2015 focus on evaluating two queries (frequent routes and  profitable cells) in real time with minimal latency and high throughput. These queries involve processing windows of thousands of records. Firstly, such processing demands efficient data structures and algorithms to minimize the processing over head. Secondly, system should partition data to process them in parallel to make system scalable.

In this paper, we present a set of data structures which can be used to evaluate two queries with O(log(n)) time complexity and a data partition technique to process events in parallel. Then we evaluate our solution within signal node as well as in distributed setups. We could process frequent routes query with 173 million dataset within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
\end{abstract}

@@

\section{Introduction}
introduction

>>>>

\maketitle
\begin{abstract}
Processing complex queries on unbounded event streams in real time, is a challenge for many organizational data processing systems. These systems are expected to process data with reduced latency to generate real time events, and high throughput to minimize the required hardware. In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real time with low latency and high throughput. These queries involve processing windows of thousands of records. Firstly, such processing demands efficient data structures and algorithms to minimize the processing overhead. Secondly, the system should partition data to process them in parallel to make it scalable.

In this paper, we present a set of data structures which can be used to evaluate the aforementioned queries with O(log(n)) time complexity and a data partitioning technique to process events in parallel. Then we evaluate our solution within signal node as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN. We could process frequent routes query with 173 million data set within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
\end{abstract}

@@

\section{Introduction}
%introduction

### 2015-03-24 paper/query_evaluation.tex
fixing typos

\textit{EventEmitter} reads the data file and generates events with a sequence number. This sequence number is used to order \textit{QueryEvaluator} results. For real implementation, each query has its own \textit{EventEmitter} which reads each line of the data file and generates an event with relevant information. \textit{QueryEvaluator} evaluates the query against the received event and generates top ten value changed event.  In general we can evaluate queries sequentially and parallel. In a sequential execution, all events goes to one \textit{QueryEvaluator} instance and in parallel execution, event stream is partitioned using the key field and evaluate simultaneously with different \textit{QueryEvaluator} instances.

@@

We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate among process instances. Figure \ref{sequential} shows such possible executions.

@@

First both processors can be executed in same JVM with same thread. In that case, both processes executes sequentially. We can make two processes parallel by executing them with two threads and communicate using a queue in same JVM or communicate using a TCP connection for different JVMs. In all these cases \textit{QueryEvaluator} processes messages sequentially and becomes the bottleneck. Next we show how to parallel \textit{QueryEvaluator} process.

\subsection{Parallel evaluation}
The general way to parallelize event processing, is to partition the events according to key and run multiple instances to process data. Therefore in this section, we explain how to parallelize the query evaluation process taking frequent rote query into consideration. Same concepts can be applied to profitable cell evaluation as well.

For frequent route query, route events can be partitioned using the hash code of the route object and process different route sets in different \textit{QueryEvaluators}. In such a setup, each \textit{QueryEvaluator} generates an event when top ten routes get changed for its route set. Since this is not the desired result, we can add another process to aggregate these top result events and generate global route change events. Figure \ref{parallel} shows the possible execution patterns with the new process.

@@

\textit{ResultAggregator} keeps a set of route counts received by \textit{QueryEvaluators}. Upon receiving a new event, it can update the existing route count values and check whether global top ten route counts has been changed. If there is a change, it can generate a top ten route changed event. This logic has the following issues and can be fixed as given bellow.

\begin{enumerate}
	\item \textbf{Expired values} : Route count of a particular route can be reduced due to expiration of some events. For an example, lets assume route count for route R1 is 10 and R1 is included in a top ten route event generated by a \textit{QueryEvaluator}. Then lets assume route count for R1 has been reduced to 7 (due to expiration events), when the next event is generated and hence R1 is not included in the next top ten route event. However since R1 value still is in the \textit{ResultAggregator} that may produce an event with this value. To avoid this problem, we send the removed route set details with the top ten route event generated by the \textit{QueryEvaluator}. In that case \textit{ResultAggregator} can first remove those routes.
	\item \textbf{Out of order events} : \textit{ResultAggregator} may receive events in a different order than those are produced at the \textit{EventEmitter}. This out of order events may change the final output of the \textit{ResultAggregator}. We order messages at the \textit{ResultAggregator} to avoid this problem. In order to order messages, we keep a message queue for each \textit{QueryEvaluator} and sent the message with least sequence number to \textit{ResultAggregator},  if a message is present in all queues. 
	\item \textbf{Intermediate top ten route events} : A route event increases the route count for that events' route. Further that can generate some expiration events which may reduce the route count for other routes. The top ten frequent routes can be changed due to these expiration events as well. Lets say in one \textit{QuerayEvaluator} case, a particular event cause two routes R1 and R2 to change and generates a top ten route change event. In the distributed setup these two routes can be in two \textit{QueryEvaluators} and that can cause two top ten route change events. These two events may result in two top ten route change events at \textit{ResultAggregator} as well instead of one. Further we observed only around 5\% of such events are generated, and system goes to the one \textit{QueryEveluator} result after all such event are being executed. Solution for such a problem may depend on the nature in which a real application, uses these results in decision making process. Therefore we could not provide any fix for this problem.
\end{enumerate}

Algorithm \ref{aggregate_algorithm} shows the algorithm for ResultAggregator with above fixes. First it removes the expired keys. Then it update the existing values or add new values and move the position of the list accordingly. Finally it generates an event if there is a change to the top list.

>>>>

\textit{EventEmitter} reads the data file and generates events with a sequence number. This sequence number is used to order \textit{QueryEvaluator} results. For real implementation, each query has its own \textit{EventEmitter} that reads each line of the data file and generates an event with the relevant information. \textit{QueryEvaluator} evaluates the query against the received event and generates top ten value changed event.  In general we can evaluate queries sequentially and in parallel. In a sequential execution, all events goes to one \textit{QueryEvaluator} instance and in a parallel execution, the event stream is partitioned using the key field and evaluate simultaneously with different \textit{QueryEvaluator} instances.

@@

We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate among process instances. Figure \ref{sequential} shows possible executions.

@@

First both processors can be executed in same JVM with same thread. In that case, both processes execute sequentially. We can make two processes parallel by executing them with two threads and communicate using a queue in same JVM or communicate using a TCP connection for different JVMs. In all these cases \textit{QueryEvaluator} processes messages sequentially and becomes the bottleneck. Next we show how to parallel \textit{QueryEvaluator} process.

\subsection{Parallel evaluation}
The general way to parallelize event processing, is to partition the events according to a key and run multiple instances to process data. Therefore in this section, we explain how to parallelize the query evaluation process taking frequent route query into consideration. Same concepts can be applied to profitable cell evaluation as well.

For frequent route query, route events can be partitioned using the hash code of the route object and process different route sets in different \textit{QueryEvaluators}. In such a setup, each \textit{QueryEvaluator} generates an event when the top ten routes get changed for its route set. Since this is not the desired result, we can add another process to aggregate these top result events and generate global route change events. Figure \ref{parallel} shows the possible execution patterns with the new process.

@@

\textit{ResultAggregator} keeps a set of route counts received by \textit{QueryEvaluators}. Upon receiving a new event, it can update the existing route count values and check whether the global top ten route counts have changed. If there is a change, it can generate a top ten route changed event. This logic has the following issues and can be fixed as given bellow.

\begin{enumerate}
	\item \textbf{Expired values} : Route count of a particular route can be reduced due to expiration of some events. For an example, lets assume route count for route R1 is 10 and R1 is included in a top ten route event generated by a \textit{QueryEvaluator}. Then lets assume route count for R1 has been reduced to 7 (due to expiration events), when the next event is generated and hence R1 is not included in the next top ten route event. However since R1 value still is in the \textit{ResultAggregator} that may produce an event with this value. To avoid this problem, we send the removed route set details with the top ten route event generated by the \textit{QueryEvaluator}. In that case \textit{ResultAggregator} can first remove those routes.
	\item \textbf{Out of order events} : \textit{ResultAggregator} may receive events in a different order than those are produced at the \textit{EventEmitter}. This out of order events may change the final output of the \textit{ResultAggregator}. We order messages at the \textit{ResultAggregator} to avoid this problem. In order to order messages, we keep a message queue for each \textit{QueryEvaluator} and sent the message with least sequence number to \textit{ResultAggregator},  if a message is present in all queues. 
	\item \textbf{Intermediate top ten route events} : A route event increases the route count for that events' route. Further that can generate some expiration events which may reduce the route count for other routes. The top ten frequent routes can be changed due to these expiration events as well. Lets say in one \textit{QuerayEvaluator} case, a particular event causes two routes R1 and R2 to change and generates a top ten route change event. In the distributed setup these two routes can be in two \textit{QueryEvaluators} and that can cause two top ten route change events. These two events may result in two top ten route change events at \textit{ResultAggregator} as well instead of one. Further we observed only around 5\% of such events are generated, and system goes to the one \textit{QueryEveluator} result after all such event are being executed. Solution for such a problem may depend on the nature in which a real application, uses these results in decision making process. Therefore we could not provide any fix for this problem.
\end{enumerate}

Algorithm \ref{aggregate_algorithm} shows the algorithm for ResultAggregator with above fixes. First it removes the expired keys. Then it updates the existing values or adds new values and moves the position in the list accordingly. Finally it generates an event if there is a change to the top list.

### 2015-03-24 paper/query_implementation.tex
fixing typos

\section{Query Implementation}
In this section, we elaborate how we have implemented the query logic to achieve O(log(n)) time complexity. The 2015 Grand Challenge problem has two queries. First is to evaluate top ten most frequent routes within last 30 minutes and second is to evaluate top ten profitable areas within last 15 minutes. We can use a composite object to keep route details for each route and fare details to each cell to compare values among routes and cells. However, each time window can have thousands of route and cell details. Therefore an efficient data structure is required to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently. Further since each event is received with route and cell details, above mentioned operations should perform using route and cell details as the key. In order to support all these functions with O(log(n)) time complexity we developed a data structure called \textit{NodeList}. Following parts of this section describes how we have achieve this time complexity with \textit{NodeList} data structure and how we have implemented the queries using that data structure.

@@

Table \ref{nodelist_api} shows the operations of this data structure. The \textit{NodeValue} is an interface to plug any user defined type as a value. For an example, the route query can use a \textit{RouteCount} object to keep route count while profitable cells query can use \textit{CellProfict} object to keep fare details. This interface contains a method called compare to define the order of the values. The \textit{NodeList} data structure provides the notion of a sorted values starting from position 0 (highest value) to its users. The \texttt{decrementPostion} operation moves the corresponding value towards position 0 until the correct position according to sorted order. Therefore this method should be called after increasing a value of an existing object. Similarly \texttt{incrementPosition} operation can be used to move a value to correct position after decreasing the value of an existing object. \texttt{GetTopValues} operation returns the top ten values of the list. Query algorithm sections further elaborate usage of these methods.

One way of implementing this data structure is to use a heap. Heap supports adding and extracting maximum value operations in O(log(n)) time. However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back. On the other hand if we use a doubly linked list to keep values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations. Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.  A \textit{Map} which points to values can be used to retrieve values using a key.

\textit{NodeList} data structure uses two data structures called \textit{OrderedList} to keep first 10 values and \textit{DynamicHeap} to keep other values. In this way it has to change only the last value of \textit{OrderedList} and maximum value of \textit{DynamicHeap} as required. Following sections describes how these data structures have been implemented with their operations.

@@

Table \ref{orderedlist_api} shows the operations of the \textit{OrderedList} data structure. First six methods has the same meanings as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object. We internally keep a counter and a pointer to tail to support these operations in O(1) time. Figure \ref{orderedlist_impl} shows the implementation details. The \textit{Map} structure (an \textit{java.util.HashMap}) is used to retrieve object pointers with O(1) expected time. Each node of the doubly linked list points to previous and next nodes while storing the \textit{NodeValue}.

@@

Table \ref{dynamicheap_api} shows the operations of the \textit{DynamicHeap} data structure. \texttt{MoveUp} and \texttt{moveDown} operations are used to move an element up in the heap if the value increased, or move down in the heap if the value decreased. Figure \ref{dynamicheap_impl} shows the implementation details. Conventional heaps only support \texttt{add} and \texttt{extractMax} operations since heap operations requires to know the element position in heap. When adding an element, we add it to last position and move that up and extracting maximum element start at array position 0. In order to support \texttt{moveUp}, \texttt{moveDown} and \texttt{remove} operations, we keep the array index with the object itself. This allows us to retrieve array index in O(1) time and move the element O(log(n)) time.

@@

Frequent route query requires to find the top frequent routes within last 30 minutes. Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.  Since queue operations are supported in O(1) time, query evaluation happens in O(log(n)) time. Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event. This algorithm assumes there is a global \textit{NodeList} object and route counts stored in an Object called \textit{RouteCount}.

@@

Processing an event can cause add, remove or update values of the existing time window objects. Therefore query processing mainly involves with performing above operations to the underlying \textit{NodeList} data structure and compare before and after top ten routes. As shown in Algorithm \ref{frequent_route_algorithm}, when an event arrives, first it puts the event to time window to keep track for last 30 minutes. Then it adds new route value or updates the existing value and moves its position. Similarly, for expired events either it removes the values or reduces the count for existing values. Finally, two lists are compared to decide whether to generate an event or not. Here we have given a rough outline of the algorithm. Real implementation considers other factors such as sequence number to order events.

@@

We have implemented profitable cells query in a similar manner to the frequent route query using  time windows. Complexity of the query raised following issues and we have solved them as given bellow.
\begin{enumerate}
	\item The profitability is calculated using the median profit and the number of empty taxis. In order to calculate median with O(log(n)) time complexity, we use two heaps (\textit{java.util.PriorityQueue}) . One is used to keep the values lesser than the median value (max heap) and other is to keep higher values (min heap). However the \texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time, since it traverse along the array to find the index of the object. We tried with our \textit{DynamicHeap} implementation to fix this. But that did not improved the performance. Firstly, these heap sizes does not grow over the value of 145 and there are same values repeating in the list increasing the probability of find a value without traversing the whole list. Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)). 
	\item In order to keep tract of empty taxis, we need to process both pickup cell and dropoff cell details for each message. One way of doing this is to send all events for same process instance and process both details with one event. However this avoid the possibility of partitioning data across different processes. We generate two events, one for pickup data and one for dropoff data to make the process parallel. 
	\item In order to handle expiration of dropoff events, we keep track of the taxi ids currently available in a particular cell. If a pickup event occurs before dropoff event expiration, then we remove that taxi id from that cell and reduce the empty taxi count. In such a case dropoff event expiration does not cause a reduction of taxi count.
\end{enumerate}

>>>>

\section{Query Implementation}
In this section, we elaborate how we have implemented the query logic to achieve O(log(n)) time complexity. The 2015 Grand Challenge problem has two queries. First is to evaluate top ten most frequent routes within the last 30 minutes and second is to evaluate top ten profitable areas within the last 15 minutes. We use a composite object to manage details for each route and fare details to each cell to compare values among routes and cells. However, each time window can have thousands of route and cell details. Therefore, an efficient data structure is required to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently. Further, since each event is received with route and cell details, the aforementioned operations should be performed using route and cell details as the key. In order to support all these functions with O(log(n)) time complexity we have developed a data structure called \textit{NodeList}. The remaining of this section describes how we have achieve this time complexity with the \textit{NodeList} data structure and how we have implemented the queries using that data structure.

@@

Table \ref{nodelist_api} shows the operations of this data structure. The \textit{NodeValue} is an interface to plug any user defined type as a value. For an example, the route query can use a \textit{RouteCount} object to keep route count while profitable cells query can use \textit{CellProfit} object to keep track of fare details. This interface contains a method called compare to define the order of the values. The \textit{NodeList} data structure provides the notion of a sorted values starting from position 0 (highest value) to its users. The \texttt{decrementPostion} operation moves the corresponding value towards position 0 until the correct position according to sorted order. Therefore this method should be called after increasing the value of an existing object. Similarly \texttt{incrementPosition} operation can be used to move a value to correct position after decreasing the value of an existing object. \texttt{GetTopValues} operation returns the top ten values of the list. The query algorithm sections further elaborate usage of these methods.

One way of implementing this data structure is to use a heap. A heap supports adding and extracting maximum value operations in O(log(n)) time. However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back. On the other hand if we use a doubly linked list to keep values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations. Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.  A \textit{Map} which points to values can be used to retrieve values using a key.

The \textit{NodeList} data structure uses two data structures called \textit{OrderedList} to keep first 10 values and \textit{DynamicHeap} to keep other values. In this way it has to change only the last value of \textit{OrderedList} and maximum value of \textit{DynamicHeap} as required. Following sections describes how these data structures have been implemented with their operations.

@@

Table \ref{orderedlist_api} shows the operations of the \textit{OrderedList} data structure. First six methods has the same meaning as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object. We internally keep a counter and a pointer to tail to support these operations in O(1) time. Figure \ref{orderedlist_impl} shows the implementation details. The \textit{Map} structure (an \textit{java.util.HashMap}) is used to retrieve object pointers with O(1) expected time. Each node of the doubly linked list points to previous and next nodes while storing the \textit{NodeValue}.

@@

Table \ref{dynamicheap_api} shows the operations of the \textit{DynamicHeap} data structure. \texttt{MoveUp} and \texttt{moveDown} operations are used to move an element up in the heap if the value increased, or move down in the heap if the value decreased. Figure \ref{dynamicheap_impl} shows the implementation details. Conventional heaps only support \texttt{add} and \texttt{extractMax} operations since heap operations require knowing the position of the element in underlying array. In order to support \texttt{moveUp}, \texttt{moveDown} and \texttt{remove} operations, we keep the array index with the object itself. As shown in figure \ref{dynamicheap_impl} \textit{HeapNode} object contains the array index and the \textit{NodeValue} given by the user. This allows us to retrieve array index in O(1) time and move the element O(log(n)) time.

@@

The frequent route query entails finding the most frequent routes within the last 30 minutes. Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.  Since queue operations are supported in O(1) time, query evaluations happen in O(log(n)) time. Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event. This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an Object called \textit{RouteCount}.

@@

Processing an event can cause add, remove or update values of the existing time window objects. Therefore query processing mainly involves performing the above operations with the underlying \textit{NodeList} data structure and comparing top ten routes before and after. As shown in Algorithm \ref{frequent_route_algorithm}, when an event arrives, first it puts the event to time window to keep track for last 30 minutes. Then it adds new route value or updates the existing value and moves its position. Similarly, for expired events either it removes the values or reduces the count for existing values. Finally, two lists are compared to decide whether to generate an event or not. Here we have given a rough outline of the algorithm. Real implementation considers other factors such as sequence number to order events.

@@

We have implemented profitable cells query in a similar manner to the frequent route query using  time windows. Complexity of the query raised following issues and we have solved them as given below.
\begin{enumerate}
	\item The profitability is calculated using the median profit and the number of empty taxis. In order to calculate median with O(log(n)) time complexity, we use two heaps (\textit{java.util.PriorityQueue}) . One is used to keep the values lesser than the median value (max heap) and other is to keep higher values (min heap). However the \texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time, since it traverse along the array to find the index of the object. We implemented a similar algorithm using our \textit{DynamicHeap} to keep higher and lower values than median with value count. In worst case it takes O(log(n)) time to add or remove a value, in best case (i.e if the value is already present and that does not cause median to change) it takes O(1) time.  But that did not improved the performance. First, these heap sizes do not grow over 145 and there are same values repeating in the list increasing the probability of finding a value without traversing the whole list. Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)). 
	\item In order to keep track of empty taxis, we need to process both pickup cell and dropoff cell details for each message. One way of doing this is to send all events for the same process instance and process both details with one event. However this avoids the possibility of partitioning data across different processes. We generate two events, one for pickup data and one for dropoff data to make the process parallel. 
	\item In order to handle expiration of dropoff events, we keep track of the taxi ids currently available in a particular cell. If a pickup event occurs before dropoff event expiration, then we remove that taxi id from that cell and reduce the empty taxi count. In such a case, dropoff event expiration does not cause a reduction of taxi count.
\end{enumerate}

### 2015-03-24 paper/related_work.tex
fixing typos

\section{Related work}
DEBS Grand Challenge problem naturally fallen into distributed stream processing domain. Many such systems have been developed with different feature sets according to the applications they use. Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track the trends in web queries. All these system are based on Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data. Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence dynamic membership management. Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets RDDs \cite{zaharia2012resilient} and Muppet \cite{lam2012muppet} provides state persistence using key value store based on Cassandra \cite{lakshman2010cassandra}.

>>>>

\section{Related work}
DEBS Grand Challenge problem naturally fallen into distributed stream processing domain. Many such systems have been developed with different feature sets according to the applications they use. Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track trends in web queries. All these systems are based on the Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data. Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence dynamic membership management. Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets RDDs \cite{zaharia2012resilient} and Muppet \cite{lam2012muppet} provides state persistence using a \textless key, value\textgreater store based on Cassandra \cite{lakshman2010cassandra}.
